Probabilistic Catalogs for Crowded Stellar Fields

### 2012-10-29 paper/ms.tex
Smaller font in tables

\newcommand{\sex}{\texttt{SExtractor }}

@@

\begin{abstract}
We introduce a probabilistic (Bayesian) method for producing catalogs
from images of crowded stellar fields. The method is capable of handling the
challenges introduced by overlapping sources and is able to determine the luminosity
function of the stars even when the precise luminosity of each star is uncertain.
This is in contrast with standard techniques which produce a single catalog,
potentially underestimating the uncertainties in any study of the stellar
population. The method is implemented using advanced Markov Chain Monte Carlo
(MCMC) techniques including Reversible Jump and Nested Sampling.
The computational feasibility of the method is
demonstrated on simulated data where the luminosity function of the stars is a
broken power-law. The parameters of the luminosity function can be recovered
with moderate uncertainty. We compare the results obtained from our method
with those obtained from the \sex software and find that the latter is
unable to correctly infer the number of stars or the parameters of the luminosity
function.
\end{abstract}

@@

\section{Introduction}
A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. {\bf The catalogs are then used for...}
Standard tools for generating catalogs include \sex
\citep{sextractor}, DAOPHOT \citep{1987PASP...99..191S}.
%\citep{2011EAS....45..351H}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Test Case & Parameter & Value \\
\hline
1 & $N$ & 100\\
  & $h_1$ & 0.3\\
  & $h_2$ & 0.6\\
  & $\alpha_1$ & 1.1\\
  & $\alpha_2$ & 2\\
  & $\sigma$ & 0.05\\
\hline
2 & $N$ & 1000\\
  & $h_1$ & 0.3\\
  & $h_2$ & 0.6\\
  & $\alpha_1$ & 1.1\\
  & $\alpha_2$ & 2\\
  & $\sigma$ & 0.3\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data.
\label{tab:truth}}
\end{table}

@@

\begin{table}
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Proposal & Notes\\
\hline
$N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from prior given
$\beta$\\
$N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars at random\\
$\beta$ & $\beta \to \beta + \delta_\beta$ & Move stars' fluxes along with
$\beta$\\
$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, put extra
term in acceptance probability \\
$(x_i,y_i)$ & $(x_i,y_i) \to (x_i,y_i)+(\delta_x, \delta_y)$ & Can move $>1$ star
in a single step \\
$f$ & $f \to f + \delta_f$ & Can move $>1$ stars' fluxes in a single step\\
\hline
\end{tabular}
\end{center}
\caption{All $\delta$ parameters are drawn from multi-scale distibutions such
that the largest steps are of order the prior width, and the smallest steps
are of order $10^{-6}$ times the prior width.\label{proposals}}
\end{table}

>>>>

\newcommand{\sex}{{\tt SExtractor}}

@@

\begin{abstract}
We introduce a probabilistic (Bayesian) method for producing catalogs
from images of crowded stellar fields. The method is capable of handling the
challenges introduced by overlapping sources and is able to determine the luminosity
function of the stars even when the precise luminosity of each star is uncertain.
This is in contrast with standard techniques which produce a single catalog,
potentially underestimating the uncertainties in any study of the stellar
population. The method is implemented using advanced Markov Chain Monte Carlo
(MCMC) techniques including Reversible Jump and Nested Sampling.
The computational feasibility of the method is
demonstrated on simulated data where the luminosity function of the stars is a
broken power-law. The parameters of the luminosity function can be recovered
with moderate uncertainty. We compare the results obtained from our method
with those obtained from the \sex~software and find that the latter is
unable to correctly infer the number of stars or the parameters of the luminosity
function.
\end{abstract}

@@

\section{Introduction}
A fundamental problem in astronomy is the construction of {\it catalogs} from
raw image data. {\bf The catalogs are then used for...}
Standard tools for generating catalogs include \sex~\citep{sextractor},
DAOPHOT \citep{1987PASP...99..191S}.
%\citep{2011EAS....45..351H}

@@

\begin{table}\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Test Case & Parameter & Value \\
\hline
1 & $N$ & 100\\
  & $h_1$ & 0.3\\
  & $h_2$ & 0.6\\
  & $\alpha_1$ & 1.1\\
  & $\alpha_2$ & 2\\
  & $\sigma$ & 0.05\\
\hline
2 & $N$ & 1000\\
  & $h_1$ & 0.3\\
  & $h_2$ & 0.6\\
  & $\alpha_1$ & 1.1\\
  & $\alpha_2$ & 2\\
  & $\sigma$ & 0.3\\
\hline
\end{tabular}
\end{center}
\caption{True parameter values used to generate the simulated data.
\label{tab:truth}}
\end{table}

@@

\begin{table}\footnotesize
\begin{center}
\begin{tabular}{|c|c|c|}
\hline
Parameter & Proposal & Notes\\
\hline
$N$ & $N \to N + \delta_N$ & Generate $\delta_N$ new stars from prior given
$\beta$\\
$N$ & $N \to N - \delta_N$ & Remove $\delta_N$ stars at random\\
$\beta$ & $\beta \to \beta + \delta_\beta$ & Move stars' fluxes along with
$\beta$\\
$\beta$ & $\beta \to \beta + \delta_\beta$ & Stars' fluxes fixed, put extra
term in acceptance probability \\
$(x_i,y_i)$ & $(x_i,y_i) \to (x_i,y_i)+(\delta_x, \delta_y)$ & Can move $>1$ star
in a single step \\
$f$ & $f \to f + \delta_f$ & Can move $>1$ stars' fluxes in a single step\\
\hline
\end{tabular}
\end{center}
\caption{All $\delta$ parameters are drawn from multi-scale distibutions such
that the largest steps are of order the prior width, and the smallest steps
are of order $10^{-6}$ times the prior width.\label{proposals}}
\end{table}
